Combinatorics of the geometry of Wilson loop diagrams II: Grassmann necklaces, dimensions, and denominators

### 2019-05-21 GWLDCII.tex
finshed reading section 4. added some exposition

Our next goal is to show that the dimension of the positroid cell defined by a Wilson loop diagram $(\cP, [n])$ has dimension $3|\cP|$.  Marcott in \hlfix{***cite this***}{highlighted for anyone skimming through} has a different proof which is geometric and more elegant, but it is not easy to track the effect of a particular propagator.  Our approach is much more explicit.

@@

We start with several lemmas, of roughly increasing degree of technicality.

@@

The next few lemmas describe how diagrams of this type are related to the corresponding diagram with propagator $p$ removed, first in terms of the Grassmann necklaces and then in terms of the Le diagrams. These technical lemmas will form the backbone of the inductive step in the main dimensionality argument.

@@

Next consider $I_n^{W}$. From the figure we see that in both cases we have $I_n^{W}(p) = n$. We are now at vertex 1, and the unassigned propagators are exactly those that appear in $V$. Therefore the algorithm continues as in $I_1^{V}$, i.e. we have $I_n^{W} = I_1^{V} \cup \{n\}$. By a similar argument, we must have $I_{n-1}^{W} = I_n^{V} \cup \{n-1\}$.

@@

Finally, if $n-2,n-1 \in I_{n-2}^{V}$ then $I_{n-2}^W(q) = n-2$ and $I_{n-2}^W(p) = n-1$ as above, but this is different to what occured in $I_{n-2}^{V}$ so we cannot use the same argument as above. We are now at vertex $n$ and only propagators $p$ and $q$ have been assigned; thus we are proceeding as in the construction of $I_n^{V}$ but without propagator $q$. By Lemma~\ref{vertex cyclic int lem} we know that $I_n^V(q) = n-5$, and from the diagram we see that the only way this could occur is if all other propagators in $V$ had already been assigned when we reached vertex $n-5$; thus $q$ was the final propagator to be assigned in $I_n^{V}$. Therefore $I_{n-2}^{W} = (I_{n}^{V} \setminus \{n-5\}) \cup \{n-1, n-2\}$ in this case. This completes all cases for $I_{n-2}^{W}$.

@@

\note{When doing the final formatting, try to ensure that Figure \ref{fig Le} and Lemma \ref{lem shape} are on the same page}
\begin{lem}\label{lem shape}
  Let $V$ and $W$ be as in Lemma~\ref{lem I}.
  The shape of the Le diagram of $V$ can be built from left to right of the following blocks: a rectangle which is 3 columns wide, one more column of the same height, and a partition shape with at most as many rows as the rectangle.
  The shape of the Le diagram of $W$ can be built from left to right of the following blocks: a rectangle with 3 columns and one more row than the first rectangle of $V$, and the same partition shape as in $V$.
\end{lem}

@@

Recall that $I_1$ determines the shape of the Le diagram. By Lemma~\ref{lem I} we know that
\newline ${n,n-1,n-2, n-3 \not\in I_1^{V}}$ (which yields the leftmost four columns of the diagram for $V$) and that $I_1^{W} = I_1^{V}\cup \{n-3\}$. This implies that the right hand boundary of the shape of $V$ is the same as the right hand boundary of the shape of $W$ except that $W$ has one additional row of 3 boxes while $V$ has an additional column in the $n-3$ position; that is, an extra column fourth from the left.
\end{proof}

As illustrated in Figure~\ref{fig Le}, the pieces of the Le diagrams of $V$ and $W$ will be called $\mathcal{A}$ and $\mathcal{B}$ in what follows. Over the course of the next few lemmas we will prove that the plusses in the $\mathcal{B}$ parts of each diagram are identical, and that the relationship between the plusses in the $\mathcal{A}$ regions can be described explicitly.

@@

\begin{proof}
  By Lemma~\ref{lem I} we have $I_{n-2}^{W}= (I_{n}^{V} \setminus \{n-5\})\cup \{n-1,n-2\}$ and $I_1^W = I_1^V\cup \{n-3\}$. We are necessarily in the case where $W$ admits a propagator $q$ (the right hand case of Figure~\ref{fig special p}) and we can see from the location of $q$ that $n-2\not\in I_{n-3}^{V}$ and that $n-5$ is the index of the lowest vertical edge in $\mathcal{B}$.  Also note that by Lemma~\ref{vertex cyclic int lem} and Corollary~\ref{no coloops} we must have $n\in I_n^V$. Thus this section of the Gramann necklace of $V$ looks like

@@

@@

  We make one further simplification: if our $W$ is in Case 2 of Figure~\ref{fig special p} and $n-1$ supports only $p$ but $n-2$ supports at least one other propagator, then flip $W$ on the line perpendicular to the edge from $n-2$ to $n-1$ to obtain a diagram with the configuration of Case 1. This eliminates the possibility that we could have $n-2 \in I_{n-2}^W$ but $n-1 \not\in I_{n-2}^W$.

@@

  From Lemma~\ref{lem shape} we know how the shapes of the Le diagrams of $V$ and $W$ are related; let $\mathcal{A}$ and $\mathcal{B}$ be as described after that lemma.  Lemmas~\ref{lem n and n-1}, \ref{lem n-2 good}, and \ref{lem n-2 bad} tell us that the three boxes of the bottom row of the Le diagram of $W$ each have a plus.  Lemmas~\ref{lem n and n-1} through \ref{lem other k} show that there is a bijection between the plusses of the Le diagram of $V$ and the plusses of the Le diagram of $W$ that are not in the bottom row.  This bijection can be described as follows:
  \begin{itemize}
  \item All plusses from $\mathcal{B}$ for $V$ maintain their positions in $\mathcal{B}$ for $W$.
  \item All plusses from the leftmost two columns (the $n$ and the $n-1$ columns) of $\mathcal{A}$ for $V$ maintain their positions in $\mathcal{A}$ for $W$.
  \item If there is a plus in the $n-2$ column of $\mathcal{A}$ in $V$ then Lemma~\ref{lem n-2 bad} applies, so there is exactly one such plus.  This plus is mapped to the $(n-5)\rightarrow (n-1)$ plus for $W$.
  \item All plusses in the $n-3$ column for $V$ are shifted one square to the left in $\mathcal{A}$ for $W$, i.e. into the $n-2$ column.
  \end{itemize}
  Note that this map is reversible and hence bijective. Indeed, the only possible ambiguity is at the $(n-5)\rightarrow (n-1)$ plus in $W$ (if it exists), which could have come from either of the first or third bullet points. However, if the third bullet point applies then by Lemma~\ref{lem n-2 bad} there is no $(n-5)\rightarrow (n-1)$ plus in $V$, i.e. the first bullet point does not apply.

>>>>

Our next goal is to show that the dimension of the positroid cell defined by a Wilson loop diagram $(\cP, [n])$ has dimension $3|\cP|$ (Theorem \ref{thm dim}).  The positroid cells associated to the Amplituhedron are of dimension $4|\cP|$. In \cite{orientation}, Marcott and the first author show that the dimension of the Deodhar component is $|\cP| +$ the dimension of associated positroid cell. In otherwords, the Deodhar component associated to a Wilson loop diagram is $4|\cP|$, showing that the geometry underlying the Wilson loop diagrams is consistent with the geometry underlying the Amplituhedron.

Marcott in \hlfix{***cite this***}{highlighted for anyone skimming through} has a different proof of the dimension of the positroid cell assicated to a Wilson loop diagram, which is geometric and more elegant, but it is not easy to track the effect of a particular propagator.  Our approach is much more explicit.

@@

We start with several lemmas, of roughly increasing degree of technicality. Lemma \ref{lem uncovered} shows that non-supporting vertices corresopond to columns of $0$s in the associated Le diagram. The proof of Theorem \ref{thm dim} depends on being able to act on Wilson loop diagrams by elements of the dihedral group to position propagators in certain positions. Lemma \ref{lem dihedral} shows that the dihedral group action does not change the dimension of of the associated Le diagram.

@@

The next few lemmas describe how diagrams of this type are related to the corresponding diagram with propagator $p$ removed, first in terms of the Grassmann necklaces (Lemma \ref{lem I}) and then in terms of the Le diagrams (Lemmas \ref{lem shape} to Lemma \ref{lem other k}) . These technical lemmas will form the backbone of the inductive step in the main dimensionality argument.

@@

Next consider $I_n^{W}$. From the figure we see that in both cases we have $I_n^{W}(p) = n$. At vertex 1 of the algorithm $I_n$, the unassigned propagators are exactly those that appear in $V$.\sanote{same invisible lemma} Therefore the algorithm continues as in $I_1^{V}$, i.e. we have $I_n^{W} = I_1^{V} \cup \{n\}$. By a similar argument, we must have $I_{n-1}^{W} = I_n^{V} \cup \{n-1\}$.

@@

Finally, if $n-2,n-1 \in I_{n-2}^{V}$ then $I_{n-2}^W(q) = n-2$ and $I_{n-2}^W(p) = n-1$ as above, but this is different to what occured in $I_{n-2}^{V}$ so we cannot use the same argument as above. When the algorithm $I_{n-2}^W$ reaches vertex $n$, the only propagators $p$ and $q$ have contributed; thus we proceed as in the construction of $I_n^{V}$ but without propagator $q$. By Lemma~\ref{vertex cyclic int lem} we know that $I_n^V(q) = n-5$, and from the diagram we see that the only way this could occur is if all other propagators in $V$ had already been assigned when we reached vertex $n-5$; thus $q$ was the final propagator to be assigned in $I_n^{V}$. Therefore $I_{n-2}^{W} = (I_{n}^{V} \setminus \{n-5\}) \cup \{n-1, n-2\}$ in this case. This completes all cases for $I_{n-2}^{W}$.

@@

\note{When doing the final formatting, try to ensure that Figure \ref{fig Le} and Lemma \ref{lem shape} are on the same page}
\begin{lem}\label{lem shape}
  Let $V$ and $W$ be as in Lemma~\ref{lem I}.
  The shape of the Le diagram of $V$ can be built from left to right of the following blocks: a rectangle which is 3 columns wide, one more column of the same height, and a \hl{partition shape}\sanote{Is this defined? shape defined by $I_i$?} with at most as many rows as the rectangle.
  The shape of the Le diagram of $W$ can be built from left to right of the following blocks: a rectangle with 3 columns and one more row than the first rectangle of $V$, and the same partition shape as in $V$.
\end{lem}

@@

Recall that $I_1$ determines the shape of the Le diagram. By the arguments in Lemma~\ref{lem I} we know that ${n,n-1,n-2, n-3 \not\in I_1^{V}}$ (which yields the leftmost four columns of the diagram for $V$) and that $I_1^{W} = I_1^{V}\cup \{n-3\}$. This implies that the right hand boundary of the shape of $V$ is the same as the right hand boundary of the shape of $W$ except that $W$ has one additional row of 3 boxes while $V$ has an additional column in the $n-3$ position; that is, an extra column fourth from the left. \sanote{Can we rephrase this in terms of row and column numbers (as in section 2) rather than use shape?}
\end{proof}

As illustrated in Figure~\ref{fig Le}, the pieces of the Le diagrams of $V$ and $W$ will be called $\mathcal{A}$ and $\mathcal{B}$ in what follows. \hlfix{Explicitly, $\mathcal(B)$ is the set of columns with labels $v < n-3$ and $\mathcal{A}$ the set of columns labeled by $n$, $n-1$ and $n-2$.}{is this accurate?} Over the course of the next few lemmas we will prove that the plusses in the $\mathcal{B}$ parts of each diagram are identical, and that the relationship between the plusses in the $\mathcal{A}$ regions can be described explicitly.

@@

\begin{proof}
  By Lemma~\ref{lem I} we have $I_{n-2}^{W}= (I_{n}^{V} \setminus \{n-5\})\cup \{n-1,n-2\}$ and $I_1^W = I_1^V\cup \{n-3\}$. We are necessarily in the case where $W$ admits a propagator $q$ (the right hand case of Figure~\ref{fig special p}) and we can see from the location of $q$ that $n-2\not\in I_{n-3}^{V}$ and that $n-5$ \hlfix{is the label of the last row in}{would like to propagate changes like this through.} $\mathcal{B}$.  Also note that by Lemma~\ref{vertex cyclic int lem} and Corollary~\ref{no coloops} we must have $n\in I_n^V$. Thus this section of the Grassmann necklace of $V$ looks like

@@

We are now ready to prove the main result of this section, showing that the dimension of the positroid cell associated to each Wilson loop diagram is $3|\cP|$.

@@

  We make one further simplification: if our $W$ is in Case 2 of Figure~\ref{fig special p} and $n-1$ supports only $p$ but $n-2$ supports at least one other propagator, then flip $W$ on the line perpendicular to the $(n-2)^{th}$ edge to obtain a diagram with the configuration of Case 1. This eliminates the possibility that we could have $n-2 \in I_{n-2}^W$ but $n-1 \not\in I_{n-2}^W$.

@@

  From Lemma~\ref{lem shape} we know how the shapes of the Le diagrams of $V$ and $W$ are related; let $\mathcal{A}$ and $\mathcal{B}$ be as described above.  Lemmas~\ref{lem n and n-1}, \ref{lem n-2 good}, and \ref{lem n-2 bad} tell us that the three boxes of the bottom row of the Le diagram of $W$ each have a plus.  Lemmas~\ref{lem n and n-1} through \ref{lem other k} show that there is a bijection between the plusses of the Le diagram of $V$ and the plusses of the Le diagram of $W$ that are not in the bottom row.  This bijection can be described as follows:
  \begin{itemize}
  \item All plusses from $\mathcal{B}$ for $V$ maintain their positions in $\mathcal{B}$ for $W$.
  \item All plusses from the leftmost two columns (the $n$ and the $n-1$ columns) of $\mathcal{A}$ for $V$ maintain their positions in $\mathcal{A}$ for $W$.
  \item If there is a plus in the $n-2$ column of $\mathcal{A}$ in $V$ then Lemma~\ref{lem n-2 bad} applies, so there is exactly one such plus.  This plus is mapped to the $(n-5)\rightarrow (n-1)$ plus for $W$.
  \item All plusses in the $n-3$ column for $V$ are shifted one square to the left in $\mathcal{A}$ for $W$, i.e. into the $n-2$ column.
  \end{itemize}
  Note that this map is reversible and hence bijective. Indeed, the only possible ambiguity is at the $(n-5)\rightarrow (n-1)$ plus in $W$ (if it exists), which could have come from either of the first or third bullet points. However, if the third bullet point applies then by Lemma~\ref{lem n-2 bad} there is no $(n-5)\rightarrow (n-1)$ plus in $V$, i.e. the first bullet point does not apply.
